Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-04-12 2017-Xudong-SMT/OpenPNets-SMT.tex
Add the LOTOS running example.
In appendix.
The description of examples is added.
TODO the analysis of the result.

The root node of the tree PN1 is the top level of the pNet
  structure. It acts as the parallel operator. There are two subnets,
  noted PN2 and PN3, each representing one of the prefix
  sub-expression. Their behaviors are synchronized in their parent PN1
  according to a set of synchronization vectors, synchronizing input
  and output action from these two subnets to give out a silent action
  $\tau$ or just making the action become visible on the top level when
  one of the subnets is working alone. The vector concrete syntax is typically
  ``<act, \_ > -> act'', in which ``\_'' means that the corresponding
  subnet is not involved in this synchronization.
  There is also a difference with the usual CCS notation of
  input/output actions: in the first vector, $!ch$ and $ch$ are
  corresponding input and output channels (see next section for a more
  complete formalisation). 
  Remark that these 4
  synchronization vectors correspond faithfully to the usual SOS
  rules of CCS operational semantics.

@@

\subsection{Open pNet with Assignments in Leaves}
\begin{figure}[h]

  \includegraphics[width=6cm]{XFIG/Enable1.eps}
%  \hspace{0.5cm}
  \includegraphics[width=6cm]{XFIG/Enable2.eps}
  \caption{Two pNet encodings for the Lotos Enable operator P>>Q }  \label{schema:enable-pnets}

\end{figure}

@@

In fact we give two encodings of this operator: the first one is
``state oriented'', in the sense that the controller has two states,
indicating where the control point is in the P>>Q process expression,
either in P or in Q. The transition between these occurs when process
P performs a $\delta(x)$ action, transferring the control to Q, and
sending a value $x$ at the same time. Process Q must be ready to
``accept'' this value through an $acc(x)$ action. The controller C1,
and the 3 synchronization vectors of the Enable1 pNet should be easy
to understand. Note 
that the first vector, transmitting an action $a1$ of P, can only
occur if $a1$ is not a $\delta$.

@@

  \begin{minipage}{6cm}
$\begin{array}{@{}r@{}l@{}}
    ot_3  = &\openrule{\{0 \xrightarrow{a}_{C_1} 1\}, ~~a\neq l \!\land\! v_3=a}
    {\ostate{00} \xrightarrow{v_3} \ostate{10}}\\
    ot_5  = &\openrule{\{0 \xrightarrow{b}_{C_2} 1\}, ~~b\neq r \!\land\! v_5=b}
    {\ostate{10} \xrightarrow{v_5} \ostate{11}}\\
  \end{array}$
  \end{minipage}
   \hspace{2mm}
  \begin{minipage}{6cm}
$\begin{array}{@{}r@{}l@{}}
    ot_7  = &\openrule{\{1 \xrightarrow{l}_{C_1} 1\}, ~~\{\xrightarrow{hb_{17}}\}, ~~v_7=hb_{17}}
    {\ostate{10} \xrightarrow{v_7} \ostate{10}}\\
  \end{array}$
  \end{minipage}

$\begin{array}{@{}r@{}l@{}}
    ot_9  = &\openrule{\{0 \xrightarrow{b}_{C_2} 1\}, ~~\{\xrightarrow{hb_{19}}\}, ~~b=!ch_{19}(v_{19}) \!\land\! b\neq r \!\land\! hb_{19}=?ch_{19}(v_{19})\!\land\! v_9=\tau}
    {\ostate{10} \xrightarrow{v_9} \ostate{11}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_{13}  = &\openrule{\{1 \xrightarrow{l}_{C_1} 1, 1 \xrightarrow{r}_{C_2} 1\}, ~~\{\xrightarrow{hb_{113}}, \xrightarrow{hb_{213}}\}, ~~hb_{113}= !ch_{113}(v_{113}) \!\land\! hb_{213}=?ch_{113}(v_{113}) \!\land\! v_{13}=\tau}
    {\ostate{11} \xrightarrow{v_{13}} \ostate{11}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_{15}  = &\openrule{\{1 \xrightarrow{l}_{C_1} 1\}, ~~\{\xrightarrow{hb_{115}}\}, ~~v_{15}=hb_{115}}
    {\ostate{11} \xrightarrow{v_{15}} \ostate{11}}\\
  \end{array}$

@@

% \newpage
% \appendix

\end{document}

>>>>

The root node of the tree PN1 is the top level of the pNet
  structure. It acts as the parallel operator. There are two subnets,
  noted PN2 and PN3, each representing one of the prefix
  sub-expression. Their behaviors are synchronized in their parent PN1
  according to a set of synchronization vectors, synchronizing input
  and output action from these two subnets to give out a silent action
  $\tau$ or just making the action become visible on the top level when
  one of the subnets is working alone. The vector concrete syntax is typically
  ``<act, \_ > -> act'', in which ``\_'' means that the corresponding
  subnet is not involved in this synchronization.
  There is also a difference with the usual CCS notation of
  input/output actions: in the first vector, $!ch$ and $?ch$ are
  corresponding input and output channels (see next section for a more
  complete formalisation). 
  Remark that these 4
  synchronization vectors correspond faithfully to the usual SOS
  rules of CCS operational semantics.

@@

\subsection{Open pNet with Assignments in Leaves}
\begin{figure}[h]
  \includegraphics[width=6cm]{XFIG/Enable1.eps}
%  \hspace{0.5cm}
  \includegraphics[width=6cm]{XFIG/Enable2.eps}
  \caption{Two pNet encodings for the Lotos Enable operator P>>Q }  \label{schema:enable-pnets}
\end{figure}

@@

In fact we give two encodings of this operator: the first one is
``state oriented'', in the sense that the controller has two states,
indicating where the control point is in the P>>Q process expression,
either in P or in Q. The transition between these occurs when process
P performs a $\delta(x)$ action, transferring the control to Q, and
sending a value $x$ at the same time. Process Q must be ready to
``accept'' this value through an $acc(x)$ action. The controller C1,
and the 3 synchronization vectors of the Enable1 pNet should be easy
to understand. Note 
that the first vector, transmitting an action $a1$ of P, can only
occur if $a_1$ is not a $\delta$.

@@

  \begin{minipage}{6cm}
$\begin{array}{@{}r@{}l@{}}
    ot_3  = &\openrule{\{0 \xrightarrow{a}_{C_1} 1\}, ~~a\neq l \!\land\! v_3=a}
    {\ostate{00} \xrightarrow{v_3} \ostate{10}}\\
    ot_5  = &\openrule{\{0 \xrightarrow{b}_{C_2} 1\}, ~~b\neq r \!\land\! v_5=b}
    {\ostate{10} \xrightarrow{v_5} \ostate{11}}\\
  \end{array}$
  \end{minipage}
   \hspace{2mm}
  \begin{minipage}{6cm}
$\begin{array}{@{}r@{}l@{}}
    ot_7  = &\openrule{\{1 \xrightarrow{l}_{C_1} 1\}, ~~\{\xrightarrow{hb_{17}}_P\}, ~~v_7=hb_{17}}
    {\ostate{10} \xrightarrow{v_7} \ostate{10}}\\
  \end{array}$
  \end{minipage}

$\begin{array}{@{}r@{}l@{}}
    ot_9  = &\openrule{\{0 \xrightarrow{b}_{C_2} 1\}, ~~\{\xrightarrow{hb_{19}}_P\}, ~~b=!ch_{19}(v_{19}) \!\land\! b\neq r \!\land\! hb_{19}=?ch_{19}(v_{19})\!\land\! v_9=\tau}
    {\ostate{10} \xrightarrow{v_9} \ostate{11}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_{13}  = &\openrule{\{1 \xrightarrow{l}_{C_1} 1, 1 \xrightarrow{r}_{C_2} 1\}, ~~\{\xrightarrow{hb_{113}}_P, \xrightarrow{hb_{213}}_Q\}, ~~hb_{113}= !ch_{113}(v_{113}) \!\land\! hb_{213}=?ch_{113}(v_{113}) \!\land\! v_{13}=\tau}
    {\ostate{11} \xrightarrow{v_{13}} \ostate{11}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_{15}  = &\openrule{\{1 \xrightarrow{l}_{C_1} 1\}, ~~\{\xrightarrow{hb_{115}}_P\}, ~~v_{15}=hb_{115}}
    {\ostate{11} \xrightarrow{v_{15}} \ostate{11}}\\
  \end{array}$

@@

% \newpage
\appendix{
\section{LOTOS Running Example}
\QIN{
We choose the LOTOS formula $(\delta(arg)+P)>>(acc(v);Q)$, composing several operators of LOTOS language, as our running example for the open automaton computing algorithm. In past papers, we already have the encoding of LOTOS operators \cite{HMZ:PDP15,henrio:Forte2016}. In Fig.~\ref{schema:lotos-pnet}, we show the encoding of the formula using the encoding of those operators.
}

\QIN{
In fact, there are two encodings of the formula. The first one uses a controller with two states to indicate the position of the control point in the top level of pNet,  either in the left hand side $(\delta(arg)+P)$ or in the other side $(acc(v);Q)$. So it is ``state oriented''. Each side of the formula is act as a subnet of the root pNet node. Three synchronization vectors are given for synchronizing the behavior of the subnets. The first vector, transmitting an action $a_1$ of P, can only occur if a1 is not a $\delta$, the vector concrete syntax is typically ``$< l, a_1 , \_ > -> a_1$'', in which ``$\_$'' means that the corresponding subnet is not involved in this synchronization. The second one composes the action from the subnets and gives out a synchronized action $\underline{\delta(x)}$, where the underlined action stands for an internal (already synchronized) action, that cannot be further synchronized at upper levels. Synchronized action is a straightforward generalization of the notion of internal actions, that will be convenient for observing internal events during model-checking. The second vector constraints that the transition occurs only when process P performs a $\delta(x)$ action, transferring the control to Q, and sending a value x at the same time. Process Q must be ready to ``accept'' this value through an acc(x) action. The last one acts similar with the the first one while it transmits the action $a_2$ of Q. These synchronization vectors make the pNet act as the Enable operator.
}

\QIN{
The second encoding is ``data oriented'', meaning that the state of the node controller is encoded in the value of the pLTS state variable(s). Here we have a single state variable s0, initialized as 0. Its value is ``0'' when the control is on P, and ``1'' when the control is on Q. The transitions in the pLTS have the more general syntax: ``<action-expr> [guard] {sequence of assignments}''. Guards and assignments can only use the values of variables of their source state and values from input variables in the action expression; and a transition can only assign variables of its target state.
}

\QIN{
Two subnets separately act as the choice and action prefix operator, in which there is a parameterized labelled transition system (pLTS) as a controller and a hole as the process. In the node PN2, there are two synchronization vectors present that action $\delta(x)$ or action from P will be transmitted to the upper level, only once and only one of them will be activated according to C2. Then the same behavior will be continued, presented by the other two synchronization vectors. The node PN3 also has its own synchronization vectors to make it act as the prefix operator.
}

\QIN{
It should be easy to see that this pNet system encodes properly the behavioral semantics of the corresponding CCS expression. Remark that it is built in a systematic structural way, one pNet node encoding each CCS operator. And it is parameterized both at the level of actions variables (occurring within the con- trollers and the synchronization vectors), and at the level of process parameters (holes).
}

\begin{figure}[h]
  \centerline{\includegraphics[width=10cm]{XFIG/LOTOS1.eps}}
  \hspace{0.5cm}
  \centerline{\includegraphics[width=10cm]{XFIG/LOTOS2.eps}}
  \caption{Two pNet encodings for the LOTOS formula }  \label{schema:lotos-pnet}
\end{figure}

\begin{figure}[h]
  \centerline{\includegraphics[width=7cm]{XFIG/LOTOSresult}}

$\begin{array}{@{}r@{}l@{}}
    ot_1  = &\openrule{\{0 \xrightarrow{l}_{C_1} 0, 0 \xrightarrow{\delta(x)}_{C_2} 1\}, ~~\{\xrightarrow{hb_{11}}_P\}, ~~hb_{11}=a_1 \!\land\! a_1\neq \delta(x) \!\land\! v_1=a_1}
    {\ostate{000} \xrightarrow{v_1} \ostate{020}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_2  = &\openrule{\{0 \xrightarrow{\delta}_{C_1} 1, 0 \xrightarrow{delta(x)}_{C_2} 1, 0 \xrightarrow{acc(x)}_{C_3} 1\}, ~~v_2=\underline{\delta(x)}}
    {\ostate{000} \xrightarrow{v_2} \ostate{111}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_3  = &\openrule{\{0 \xrightarrow{\delta}_{C_1} 1, 0 \xrightarrow{p}_{C_2} 2, 0 \xrightarrow{acc(x)}_{C_3} 1\}, ~~\{\xrightarrow{hb_{13}}_P\}, ~~hb_{13}=\delta(x) \!\land\! v_3=\underline{\delta(x)}}
    {\ostate{000} \xrightarrow{v_3} \ostate{121}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_4  = &\openrule{\{0 \xrightarrow{l}_{C_1} 0, 2 \xrightarrow{p}_{C_2} 2\}, ~~\{\xrightarrow{hb_{14}}_P\}, ~~hb_{14}=a_1 \!\land\! a_1\neq \delta(x) \!\land\! v_4=a_1}
    {\ostate{020} \xrightarrow{v_4} \ostate{020}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_5  = &\openrule{\{0 \xrightarrow{\delta}_{C_1} 1, 2 \xrightarrow{p}_{C_2} 2, 0 \xrightarrow{acc(x)}_{C_3} 1\}, ~~\{\xrightarrow{hb_{15}}_P\}, ~~hb_{15}=\delta(x) \!\land\! v_5=\underline{\delta(x)}}
    {\ostate{020} \xrightarrow{v_5} \ostate{121}}\\
  \end{array}$

$\begin{array}{@{}r@{}l@{}}
    ot_6  = &\openrule{\{1 \xrightarrow{r}_{C_1} 1, 1 \xrightarrow{q}_{C_3} 1\}, ~~\{\xrightarrow{hb_{16}}_Q\}, ~~hb_{16}=a_2 \!\land\! a_2\neq acc(x) \!\land\! v_6=a_2}
    {\ostate{111} \xrightarrow{v_6} \ostate{111}}\\
  \end{array}$

  $\begin{array}{@{}r@{}l@{}}
    ot_7  = &\openrule{\{1 \xrightarrow{r}_{C_1} 1, 1 \xrightarrow{q}_{C_3} 1\}, ~~\{\xrightarrow{hb_{17}}_Q\}, ~~hb_{17}=a_2 \!\land\! a_2\neq acc(x) \!\land\! v_7=a_2}
    {\ostate{121} \xrightarrow{v_7} \ostate{121}}\\
  \end{array}$

\caption{The open automaton of the LOTOS formula ($\delta$(arg)+P)>>(acc(v);Q)}  \label{schema:lotos-result}
\end{figure}
}
\end{document}
